# 01_Create Raw Seurat Object
Loading raw data. joint_bcs as intersect of gex and hto gives the same result as gex and adt. Keeping it as it is.

```{r 01_libraries}
suppressPackageStartupMessages({
  library(Seurat)
  library(DropletUtils)
  library(ggplot2)
  library(Matrix)
  library(tidyverse)
  library(clustree)
  library(reticulate) # load packages required by leiden algo
  reticulate::use_condaenv("/mnt/sda4/conda/envs/R")
  library(viridis) # colors
  library(DoubletFinder)
  options(future.globals.maxSize = 3.0 * 1e9) # 3GB
  library(sceasy) # convert to anndata, scvi integration
  library(dittoSeq)
  library(dsb)
})
```

## stuff
```{r 01_functions}
set.seed(123)

source("/mnt/sda4/singleCell_LAB/HitGlio_sample1-test/scripts/functions.R")
source("/mnt/sda4/singleCell_LAB/HitGlio_sample1-test/scripts/00_Global_vars.R")

sample_path <- file.path("/mnt/sdb1/runs/sample1_multilane_spec_index_NNN/outs/multi/count/raw_feature_bc_matrix") # nolint
```

## create sparse matrix
```{r, warning = FALSE}
sparse_matrix_list <- seu_create_sparse_matrix_list(sample_path)
```
## filter gex with DropletUtils
```{r}
sparse_matrix_filt <- filter_gex_matrix_dropletutils(sparse_matrix_list)
```
## Create Raw Seurat Object
```{r}
seu_raw_object <- create_seurat_object(sparse_matrix_filt)
```
## Demultiplex Raw Seurat Object
```{r}
seu_raw_object_demux <- seu_demultiplex(seu_raw_object)
```
## Subset singlets and negatives
```{r}
seu_singlet_negative_list <- subset_singlets(seu_raw_object_demux)
```
### Inspect Singlets and negatives
```{r}
seu_singlet_negative_list$seu_singlet
seu_singlet_negative_list$seu_negative
```
## Add quality features and cycle scroring
```{r}
seu_singlet_b4_q_filtering <- add_quality_features(
  seu_singlet_negative_list$seu_singlet
) |>
  add_cell_cycle_scoring()
```

# 02_Check Quality of the raw data
## Normalize
```{r}
seu_singlet_b4_q_filt_norm <- seu_normalize_var_scale_pca(
  seu_singlet_b4_q_filtering,
  normalize = "log",
  cluster.name = "cluster.log",
  pca.reduction.name = "pca",
  umap.reduction.name = "umap.log",
  dims = 1:30,
  k.param = 20,
  algorithm = 4,
  resolution = 0.5,
  group.singletons = TRUE
)
```
## Visualize
```{r}
viz_quality(
  seu_singlet_b4_q_filt_norm,
  assay = "RNA", q_features,
  reduction = "umap.log", gr_by_features
)
plot_scatter(seu_singlet_b4_q_filt_norm)
ElbowPlot(seu_singlet_b4_q_filt_norm, ndims = 40)
```

# 03_Subset cells based on quality assesment
## Subset
```{r}
seu_singlet_q_filt_norm <- subset(
  seu_singlet_b4_q_filt_norm,
  subset = nFeature_RNA > 200 &
    nFeature_RNA < 7000 &
    nCount_RNA < 50000 &
    nCount_RNA > 500 &
    percent_mito < 8 &
    nCount_ADT < quantile(seu_singlet_b4_q_filt_norm$nCount_ADT, 0.99)
)
```
## Visualize
```{r}
viz_quality(
  seu_singlet_q_filt_norm,
  assay = "RNA", q_features,
  reduction = "umap.log", gr_by_features
)
plot_scatter(seu_singlet_q_filt_norm)
```

# 04_Remove Doublets Using DoubletFinder
## normalize after quality subsetting (SCT)
```{r, warning = FALSE}
seu_singlet <- seu_normalize_var_scale_pca(
  seu_singlet_q_filt_norm,
  normalize = "sct",
  cluster.name = "cluster.sct",
  pca.reduction.name = "pca.sct",
  umap.reduction.name = "umap.sct",
  dims = 1:10, # change to 10 in case of removing doublets
  k.param = 20,
  algorithm = 4,
  resolution = 0.2, # res probably could be lower, like 0.2
  group.singletons = TRUE
)
DefaultAssay(seu_singlet) <- "SCT"
```
## Remove Doublets
```{r}
seu_singlet <- remove_doublets_w_dfinder(seu_singlet)
```
## subset tru singlets
```{r}
seu_singlet <- subset(
  seu_singlet,
  subset = DF.classifications_0.25_0.09_1657 == "Singlet" # generalize it!
)
```
## GEX
### Normalize and ReCluster
```{r, warning = FALSE}
seu_singlet <- seu_normalize_var_scale_pca(
  seu_singlet,
  normalize = "sct",
  cluster.name = "cluster.sct",
  pca.reduction.name = "pca.sct",
  umap.reduction.name = "umap.sct",
  dims = 1:20,
  k.param = 30,
  algorithm = 4,
  resolution = 0.1,
  group.singletons = FALSE
)
```
### Visualize
```{r}
viz_quality(
  seu_singlet,
  assay = "SCT", t_quality,
  reduction = "umap.sct", gr_by_features_sct
)
seu_dittoDotPlot(
  seu_singlet,
  vars = t_quality,
  group.by = "cluster.sct"
)
```
## ADT
### Normalize using CLR
only for reference, DSB normalization in the main one
```{r}
DefaultAssay(seu_singlet) <- "ADT"
seu_singlet <- NormalizeData(
  seu_singlet,
  assay = "ADT",
  normalization.method = "CLR"
)
```
### Visualize
```{r}
viz_quality(
  seu_singlet,
  assay = "ADT", c(t_adt_quality, isotype_controls),
  reduction = "umap.sct", gr_by_features_sct
)
seu_dittoDotPlot(
  seu_singlet,
  vars = t_adt_quality,
  group.by = "cluster.sct"
)
```

# Remove cd8+ from myeloid clusters
## Subset myeloid cluster to filter out cd8 cells -------------------
```{r subset myelo}
DefaultAssay(seu_singlet) <- "SCT"
Idents(seu_singlet) <- "cluster.sct"

seu_singlet_myelo <- subset(seu_singlet, idents = c(2, 6, 10, 11))
DefaultAssay(seu_singlet_myelo) <- "SCT"
seu_singlet_myelo
Idents(seu_singlet_myelo)

seu_singlet_myelo <- seu_normalize_var_scale_pca(
  seu_singlet_myelo,
  normalize = "sct",
  cluster.name = "cluster.sct.myelo",
  run_pca = TRUE,
  pca.reduction.name = "pca.myelo.sct",
  umap.reduction.name = "umap.myelo.pca.sct",
  dims = 1:5,
  k.param = 10,
  algorithm = 4,
  resolution = 1.2,
  group.singletons = FALSE
)
```

```{r viz isolated myelo clusters}
viz_quality(seu_singlet_myelo,
  assay = "SCT", q_features = t_quality,
  reduction = "umap.myelo.pca.sct", gr_by_features = gr_by_features_sct_myelo
)

seu_dittoDotPlot(
  seu_singlet_myelo,
  vars = vars,
  group.by = "cluster.sct.myelo"
)

DefaultAssay(seu_singlet) <- "ADT"
viz_quality(seu_singlet_myelo,
  assay = "ADT", q_features = t_adt_quality,
  reduction = "umap.myelo.pca.sct", gr_by_features = NULL
)
```

# Remove selected cells from seu_singlet (original) clusters
```{r remove cells}
Idents(seu_singlet_myelo) <- "cluster.sct.myelo"
clusters_to_remove <- subset(seu_singlet_myelo, idents = c(11, 19, 25, 28))
clusters_to_remove

seu_singlet <- subset(
  seu_singlet,
  cells = setdiff(Cells(seu_singlet), Cells(clusters_to_remove))
)
seu_singlet
```

# Rereun normalize after removing CD8 cells from myelo clusters ------------
```{r normalize after removing cd8 from myelo clusters}
seu_singlet <- seu_normalize_var_scale_pca(
  seu_singlet,
  normalize = "sct",
  cluster.name = "cluster.sct",
  run_pca = TRUE,
  pca.reduction.name = "pca.sct",
  umap.reduction.name = "umap.sct",
  dims = 1:30,
  k.param = 20,
  algorithm = 4,
  resolution = 0.5,
  group.singletons = TRUE
)
seu_singlet <- NormalizeData(
  seu_singlet,
  assay = "ADT",
  normalization.method = "CLR"
)
```

```{r viz seu_singlet after removing cd8 from myelo clusters}
viz_quality(seu_singlet,
  assay = "SCT", q_features = t_quality,
  reduction = "umap.sct", gr_by_features = gr_by_features_sct
)

seu_dittoDotPlot(
  seu_singlet,
  vars = vars,
  group.by = "cluster.sct"
)

DefaultAssay(seu_singlet) <- "ADT"
viz_quality(seu_singlet,
  assay = "ADT", q_features = t_adt_quality,
  reduction = "umap.sct", gr_by_features = NULL
)
```


# 06_DSB normalization
Move here from 05_DSB (and create functions which should go to functions.R)

# 07_Remove proteins that do not pass DSB thresholds
remove also isotype-controld
## Strong YES
### Features to remove
```{r}
features_to_remove <- paste0("TotalSeqC-", c(
  "CD279", "CD161", "CD274", "CD152", "CD223", "CD366", "CD357",
  "CD80", "CD86", "CD197", "CD184", "CD335", "CD106", "CD314", "TMEM119",
  "CD137", "CD28",
  "IgG2b-Ctrl", "IgG2a-Ctrl", "IgG-Ctrl", "IgG1-Ctrl"
))
features_to_remove
valid_features <- intersect(features_to_remove, rownames(seu_singlet))
```

### Move counts from adt to adt.dsb.threshold
(ast.dsb.treshold contains only dsb-normalized data)
```{r}
seu_singlet[["ADT.dsb.threshold"]]@counts <- seu_singlet[["ADT"]]@counts
```

### Remove features from both 'counts' and 'data'
```{r}
assay <- seu_singlet[["ADT.dsb.threshold"]]
assay@counts <- assay@counts[!rownames(assay@counts) %in% valid_features, ]
assay@data <- assay@data[!rownames(assay@data) %in% valid_features, ]
assay@meta.features <- assay@meta.features[
  !rownames(assay@meta.features) %in% valid_features,
]
```

### Update the assay in the Seurat object
```{r}
seu_singlet[["ADT.dsb.threshold"]] <- assay
seu_singlet <- Seurat::UpdateSeuratObject(seu_singlet)
```

### Validate
```{r}
dim(seu_singlet[["ADT.dsb.threshold"]])
dim(seu_singlet[["ADT"]])
dim(seu_singlet[["SCT"]])
```

# should I also repeat SCT normalization here after removing adt features?
No, for now, but unsure?


# Export counts from SCT and ADT.dsb.threshold for totalVI
```{r, eval=FALSE}
seu_singlet
adata <- convertFormat(
  seu_singlet,
  from = "seurat",
  to = "anndata",
  main_layer = "counts",
  # assay = "SCT",
  assay = "ADT.dsb.threshold",
  drop_single_values = FALSE,
  outFile = "seu_singlet_adt_09-jan-25.h5ad"
)
adata
```

# After totalVI run, bring back calculated latent
```{r Go to python and bring back latent.csv}
latent <- read.csv("/mnt/sda4/singleCell_LAB/HitGlio_sample1-test/latents/latent-seu_singlet_09-jan-24.csv")

latent_mtx <- as.matrix(latent)
rownames(latent_mtx) <- colnames(seu_singlet)
latent_mtx <- latent_mtx[, -1] # remove 1st col
dim(latent_mtx)
colnames(latent_mtx) <- paste0("totalvi_", 1:20)
colnames(latent_mtx)

DefaultAssay(seu_singlet) <- "SCT"
seu_singlet
colnames(seu_singlet@meta.data)

seu_singlet[["totalvi.sct"]] <- CreateDimReducObject(
  embeddings = latent_mtx,
  key = "totalvi_",
  assay = DefaultAssay(seu_singlet)
)
seu_singlet
colnames(seu_singlet@meta.data)
```

# Clustering po dodaniu latent
nie normalizuje bo nie trzeba i nie robie pca poniewaz używam totalvi
```{r clustering after adding latent}
seu_singlet <- seu_normalize_var_scale_pca(
  seu_singlet,
  normalize = FALSE,
  cluster.name = "cluster.totalvi",
  run_pca = FALSE, # not running pca, changing clusters in totalvi reduction
  pca.reduction.name = "totalvi.sct", # instead of pca
  umap.reduction.name = "umap.totalvi.sct",
  dims = 1:20,
  k.param = 40,
  algorithm = 4,
  resolution = 1.4,
  group.singletons = FALSE
)
```

## Visualize
```{r}
viz_quality(
  seu_singlet,
  assay = "SCT", c(t_quality),
  reduction = "umap.totalvi.sct", gr_by_features_totalvi
)
seu_dittoDotPlot(
  seu_singlet,
  vars = t_quality,
  group.by = "cluster.totalvi"
)
```
```{r}
Features(seu_singlet)

DefaultAssay(seu_singlet) <- "ADT.dsb.threshold"

viz_quality(
  seu_singlet,
  assay = "ADT.dsb.threshold", c(t_adt_quality),
  reduction = "umap.totalvi.sct", gr_by_features_totalvi
)
seu_dittoDotPlot(
  seu_singlet,
  vars = t_adt_quality,
  group.by = "cluster.totalvi"
)
```

# 18. Vizualize ADT features after all that cleanup
```{r}
DefaultAssay(seu_singlet) <- "ADT.dsb.threshold"
features <- get_features(seu_singlet, "ADT.dsb.threshold")
length(features)
f1 <- features[1:10]
f2 <- features[11:21]
f_list <- list(f1, f2)

for (i in seq_along(f_list)) {
  seu_FeaturePlot(
    seu_singlet, "ADT.dsb.threshold",
    reduction = "umap.totalvi.sct",
    color = inferno,
    features = f_list[[i]],
    max.cutoff = NA,
    show = FALSE,
    save_path = paste0("singleCell_LAB/HitGlio_sample1-test/plots/09-jan-25_sct_dsb-threshold/panel_", i), # nolint
    ggheight = 20,
    ggwidth = 20
  )
}
```

# Heatmap of top 5 genes in all clusters
```{r}
DefaultAssay(seu_singlet) <- "SCT"
Idents(seu_singlet) <- "cluster.totalvi"

top_genes <- Seurat::FindAllMarkers(
  seu_singlet,
  test_use = "MAST", only.pos = TRUE
) |>
  group_by(cluster) |>
  top_n(n = 5, wt = avg_log2FC)

DoHeatmap(seu_singlet, features = unique(top_genes$gene)) +
  scale_fill_viridis_c()
```

?DoHeatmap






















```{r}
cluster_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 1,
  min.pct = 0.25
)

cluster_markers_tb <- as_tibble(cluster_markers, rownames = "gene") |>
  select(gene, avg_log2FC) |>
  remove_rownames() |>
  column_to_rownames("gene")


pheatmap::pheatmap(t(cluster_markers_tb[1:20, ]),
  cluster_rows = FALSE, cluster_cols = FALSE,
  scale = "row",
  color = viridis::viridis(25, option = "B"),
  border_color = NA
)


cluster_markers_tb[1:20, ]

rownames(cluster_markers[1:20, ])
cluster_markers[1:20, 2]
cluster_markers[1:20, ]
```

