---
title: "HIT-GLIO sample1 analysis #1"
author: "MF"
date: "`r Sys.Date()`"
---

according to: https://satijalab.org/seurat/articles/hashing_vignette

httpgd::hgd_url()

rmarkdown::render("/mnt/sda4/singleCell_LAB/HitGlio_sample1-test/scripts/01_demux_filter.Rmd")


Analysis:
1. Load cellranger filtered data
2. Demux
3. Subset singlets
4. Filter
5. Remove Doublets
6. totalVI
7. Analyze totalVI results
8. Seubset Myeloid clusters and filter CD8 cells

Ale inna analiza może wyglądać inaczej:
1. Load cellranger raw data
2. DSB
3. Doublets
4. etc
(w tym przypadku robię branch i modyfikuję analizę)

Ale też może zdażyć się coś takiego, że będę chciał modyfikować niektóre parametry w analizie którą aktualnie uważam za main. W tym przypadku nie chcę liczyć wszystkiego od początku za każdym razem.
(w tym przypadku uruchamiam chunksy samodzielnie, a jak już coś ustalę, w porównaniu z wynikami poprzednimi, które mam już w html, puszczam knit i jestem w domku.)

# Load Libraries
```{r Libraries, warning = FALSE, message = FALSE}
suppressPackageStartupMessages({
  library(Seurat)
  library(DropletUtils)
  library(ggplot2)
  library(Matrix)
  library(tidyverse)
  library(clustree)
  library(reticulate) # load packages required by leiden algo
  reticulate::use_condaenv("/mnt/sda4/conda/envs/R")
  library(viridis) # colors
  library(DoubletFinder)
  options(future.globals.maxSize = 3.0 * 1e9) # 3GB
  library(sceasy) # convert to anndata, scvi integration
  library(dittoSeq)
  library(dsb)
})
```

# Various
```{r Paths, warning = FALSE, message = FALSE}
source("/mnt/sda4/singleCell_LAB/HitGlio_sample1-test/scripts/functions.R")

set.seed(123)

sample_path <- file.path("/mnt/sdb1/runs/sample1_multilane_spec_index_NNN/outs/multi/count/raw_feature_bc_matrix") # nolint
print(sample_path)

s_genes <- cc.genes$s.genes
g2m_genes <- cc.genes$g2m.genes

inferno <- viridis::scale_color_viridis(option = "inferno")
```

# Initial steps (load 10X)
Loading raw data. joint_bcs as intersect of gex and hto gives the same result as gex and adt. Keeping it as it is.
```{r Initial steps}
sparse_matrix <- Seurat::Read10X(data.dir = sample_path)
gex <- sparse_matrix$`Gene Expression`
adt <- sparse_matrix$`Antibody Capture`[grepl(
  "TotalSeqC", rownames(sparse_matrix$`Antibody Capture`)
), ]
hto <- sparse_matrix$`Antibody Capture`[grepl(
  "Hashtag", rownames(sparse_matrix$`Antibody Capture`)
), ]
rownames(hto)
rownames(hto) <- c("tumor", "csf")

joint_bcs <- intersect(colnames(gex), colnames(hto))
length(joint_bcs)

gex <- gex[, joint_bcs]
adt <- adt[, joint_bcs]

hto <- as.matrix(hto[, joint_bcs]) # not sure if now or later
```

# Identify empty cells with DropletUtils
```{r dropletutils}
e_out <- emptyDrops(gex)

is_cell <- e_out$FDR <= 0.01
table(is_cell)

is_cell[is.na(is_cell)] <- FALSE

gex_filt <- gex[, is_cell]

dim(gex_filt)
```

# create seurat object
```{r create object}
seu <- Seurat::CreateSeuratObject(
  counts = Matrix::Matrix(as.matrix(gex_filt), sparse = TRUE)
)
seu <- Seurat::NormalizeData(seu)
seu <- Seurat::FindVariableFeatures(seu, selection.method = "mean.var.plot")
seu <- Seurat::ScaleData(seu, features = Seurat::VariableFeatures(seu))

stained_cells <- colnames(gex_filt)
hto_stained <- hto[, stained_cells]
adt_stained <- adt[, stained_cells]

seu[["HTO"]] <- Seurat::CreateAssayObject(counts = hto_stained)
seu[["ADT"]] <- Seurat::CreateAssayObject(counts = adt_stained)

seu <- Seurat::NormalizeData(seu, assay = "HTO", normalization.method = "CLR")
seu <- Seurat::NormalizeData(seu, assay = "ADT", normalization.method = "CLR")
```

# Demultiplex
```{r demux}
seu <- Seurat::MULTIseqDemux(
  seu,
  assay = "HTO",
  quantile = 0.7,
  autoThresh = TRUE,
  maxiter = 5,
  qrange = seq(from = 0.1, to = 0.9, by = 0.05),
  verbose = TRUE
)
table(seu$MULTI_ID)
# tumor 15768
# csf 1701
```

# Subset singlets
```{r Subset singlets}
seu_singlet <- subset(seu, idents = c("tumor", "csf"))
seu_singlet
```

```{r Add quality features to singets}
seu_singlet <- add_quality_features(seu_singlet)

seu_singlet <- CellCycleScoring(
  seu_singlet,
  s.features = s_genes,
  g2m.features = g2m_genes,
  set.ident = TRUE
)

colnames(seu_singlet@meta.data)
```

# Normalize before quality filtering (check how it looks like)
może od razu SCT?
"_temp" bo nie bede potrzebowal tego pozniej
```{r Normalize, warning = FALSE}
seu_singlet_q_filt_temp <- seu_normalize_var_scale_pca(
  seu_singlet,
  normalize = "log",
  cluster.name = "cluster.log",
  pca.reduction.name = "pca",
  umap.reduction.name = "umap.log",
  dims = 1:30,
  k.param = 20,
  algorithm = 4,
  resolution = 0.5,
  group.singletons = TRUE
)
seu_singlet_q_filt_temp
colnames(seu_singlet_q_filt_temp@meta.data)

ElbowPlot(seu_singlet_q_filt_temp, ndims = 40)
```

```{r Visualize before quality filtering}
viz_quality <- function(
    seu, assay, q_features = NULL,
    reduction, gr_by_features = NULL) {
  #
  purrr::walk(gr_by_features, ~ seu_DimPlot(
    seu,
    reduction = reduction,
    group.by = .x
  ) |> print())

  purrr::walk(q_features, ~ seu_FeaturePlot(
    seu,
    assay = assay,
    features = .x,
    reduction = reduction,
    color = inferno,
    label = TRUE,
    max.cutoff = NA
  ) |> print())

  purrr::walk(q_features, ~ seu_violin_plot(seu, features = .x) |> print())
}

q_features <- c(
  "nCount_RNA",
  "nFeature_RNA",
  "percent_mito",
  "percent_ribo",
  "percent_hemoglob"
)
gr_by_features <- c("MULTI_ID", "cluster.log")

viz_quality(
  seu_singlet_q_filt_temp,
  assay = "RNA", q_features,
  reduction = "umap.log", gr_by_features
)

plot_scatter <- function(data) {
  FeatureScatter(
    data,
    "nFeature_RNA",
    "nCount_ADT"
  )
}
plot_scatter(seu_singlet)
```

# Subset based on quality assesment
```{r, Subset cells based on quality assesment}
seu_singlet <- subset(
  seu_singlet,
  subset = nFeature_RNA > 200 &
    nFeature_RNA < 7000 &
    nCount_RNA < 50000 &
    nCount_RNA > 500 &
    percent_mito < 8
)
```

subset ADT quantile 0.99
```{r subset adt quality}
seu_singlet <- subset(
  seu_singlet,
  subset = nCount_ADT < quantile(seu_singlet$nCount_ADT, 0.99)
)
plot_scatter(seu_singlet)

seu_singlet
```

# Remove doublets
```{r Normalize after q filtering for DOUBLET REMOVAL}
seu_singlet <- seu_normalize_var_scale_pca(
  seu_singlet,
  normalize = "sct",
  cluster.name = "cluster.sct",
  pca.reduction.name = "pca.sct",
  umap.reduction.name = "umap.sct",
  dims = 1:10, # change to 10 in case of removing doublets
  k.param = 20,
  algorithm = 4,
  resolution = 0.2, # res probably could be lower, like 0.2
  group.singletons = TRUE
)
# seu_singlet <- NormalizeData(
#   seu_singlet,
#   assay = "ADT.dsb",
#   normalization.method = "CLR"
# )
seu_singlet
```

```{r Remove doublets based on SCT initial clustering}
homotypic_prop <- modelHomotypic(seu_singlet$cluster.sct)
nExp_poi <- round(0.075 * nrow(seu_singlet)) # estimate ~7.5% doublets
nExp_poi_adj <- round(nExp_poi * (1 - homotypic_prop))

seu_singlet_temp <- DoubletFinder::doubletFinder(
  seu_singlet,
  PCs = 1:10,
  pN = 0.25,
  pK = 0.09,
  nExp = nExp_poi_adj,
  reuse.pANN = FALSE,
  sct = TRUE
)
seu_singlet_temp
colnames(seu_singlet_temp@meta.data)
head(seu_singlet_temp@meta.data)

seu_singlet <- subset(
  seu_singlet_temp,
  subset = DF.classifications_0.25_0.09_1651 == "Singlet"
)

DefaultAssay(seu_singlet) <- "SCT"
seu_singlet <- seu_normalize_var_scale_pca(
  seu_singlet,
  normalize = "sct",
  cluster.name = "cluster.sct",
  pca.reduction.name = "pca.sct",
  umap.reduction.name = "umap.sct",
  dims = 1:20,
  k.param = 30,
  algorithm = 4,
  resolution = 0.8,
  group.singletons = FALSE
)

DefaultAssay(seu_singlet) <- "SCT"
seu_singlet
```

# Viz. after quality subset and doublet removal
gr_by_features and reduction changed to .sct
```{r viz. after q filtering}
gr_by_features <- c("MULTI_ID", "cluster.sct")

viz_quality(
  seu_singlet,
  assay = "SCT", q_features,
  reduction = "umap.sct", gr_by_features
)
```

# DSB
```{r subset negative/positive cells}
# subset empty drop/background and cells
seu_negative <- subset(
  seu,
  idents = "Negative"
)
seu_negative

neg_filt <- GetAssayData(
  seu_negative,
  layer = "counts", assay = "ADT"
) |> as.matrix()

adt_filt <- GetAssayData(
  seu_singlet,
  layer = "counts", assay = "ADT"
) |> as.matrix()

dim(neg_filt)
dim(adt_filt)
```

```{r Run DSB}
rownames(adt_filt)

isotype_controls <- c("TotalSeqC-IgG2b-Ctrl", "TotalSeqC-IgG2a-Ctrl", "TotalSeqC-IgG-Ctrl", "TotalSeqC-IgG1-Ctrl")

# normalize and denoise with dsb with
cells_dsb_norm <- DSBNormalizeProtein(
  cell_protein_matrix = adt_filt,
  empty_drop_matrix = neg_filt,
  denoise.counts = TRUE,
  use.isotype.control = TRUE,
  isotype.control.name.vec = isotype_controls
)

dim(cells_dsb_norm)

seu_singlet[["ADT.dsb"]] <- CreateAssayObject(data = cells_dsb_norm)
```

# INSPECT features (mainly for CD8) after doublet removal for SCT clustering 
```{r, inspect cd8 after doublet removal}
DefaultAssay(seu_singlet) <- "SCT"

vars <- c("CD8A", "CD8B", "CD4")
gr_by_features <- c("MULTI_ID", "cluster.sct")

viz_quality(seu_singlet,
  assay = "SCT", q_features = vars,
  reduction = "umap.sct", gr_by_features = gr_by_features
)

seu_dittoDotPlot(
  seu_singlet,
  vars = vars,
  group.by = "cluster.sct"
)

DefaultAssay(seu_singlet) <- "ADT.dsb"
adt_features <- c("TotalSeqC-CD8", "TotalSeqC-CD4", isotype_controls)
isotype_controls <- c("TotalSeqC-IgG2b-Ctrl", "TotalSeqC-IgG2a-Ctrl", "TotalSeqC-IgG-Ctrl", "TotalSeqC-IgG1-Ctrl")

viz_quality(seu_singlet,
  assay = "ADT.dsb", q_features = c(adt_features, isotype_controls),
  reduction = "umap.sct", gr_by_features = NULL
)
```

```{r testy}
viz_quality(seu_singlet,
  assay = "ADT.dsb", q_features = "TotalSeqC-CD52",
  reduction = "umap.sct", gr_by_features = NULL
)
```


# 18. Vizualize ADT features after all that cleanup -------------
```{r}
DefaultAssay(seu_singlet) <- "ADT.dsb"
features <- Features(seu_singlet)
length(features)
f1 <- features[1:10]
f2 <- features[11:21]
f3 <- features[22:32]
f4 <- features[33:42]
f_list <- list(f1, f2, f3, f4)

for (i in seq_along(1:4)) {
  seu_FeaturePlot(
    seu_singlet, "ADT.dsb",
    reduction = "umap.sct",
    color = inferno,
    features = f_list[[i]],
    max.cutoff = NA,
    show = FALSE,
    save_path = paste0("plots/18-dec-24_sct_dsb_2/adt-panel_", i),
    ggheight = 20,
    ggwidth = 20
  )
}
```

```{r}
s <- Seurat::FindNeighbors(
  object = seu_singlet, dims = NULL, assay = "ADT.dsb",
  features = features, k.param = 30,
  verbose = FALSE
)

# direct graph clustering
s <- Seurat::FindClusters(
  object = s, resolution = 1,
  algorithm = 3,
  graph.name = "ADT.dsb_snn",
  verbose = FALSE
)


d <- cbind(
  s@meta.data,
  as.data.frame(t(s@assays$ADT.dsb@data))
  # s@reductions$umap@cell.embeddings)
)

library(magrittr)
adt_plot <- d %>%
  dplyr::group_by(ADT.dsb_snn_res.1) %>%
  dplyr::summarize_at(.vars = features, .funs = median) %>%
  tibble::remove_rownames() %>%
  tibble::column_to_rownames("ADT.dsb_snn_res.1")
# plot a heatmap of the average dsb normalized values for each cluster
pheatmap::pheatmap(t(adt_plot),
  color = viridis::viridis(25, option = "B"),
  fontsize_row = 8, border_color = NA
)
```

# Set Thresholds Using Isotype Controls
should be set for each individual isotype
```{r}
# Negative control proteins
neg_controls <- isotype_controls
neg_values <- cells_dsb_norm[neg_controls, ]

# Maximum noise level
threshold <- max(neg_values)

print(threshold) # Use this value as a guide for positivity threshold

cells_dsb_norm[cells_dsb_norm < threshold] <- 0 # no expression below threshold

seu_singlet[["ADT.dsb.threshold"]] <- CreateAssayObject(data = cells_dsb_norm)

DefaultAssay(seu_singlet) <- "ADT.dsb.threshold"
features <- Features(seu_singlet)
length(features)
f1 <- features[1:10]
f2 <- features[11:21]
f3 <- features[22:32]
f4 <- features[33:42]
f_list <- list(f1, f2, f3, f4)

for (i in seq_along(1:4)) {
  seu_FeaturePlot(
    seu_singlet, "ADT.dsb.threshold",
    reduction = "umap.sct",
    color = inferno,
    features = f_list[[i]],
    max.cutoff = NA,
    show = FALSE,
    save_path = paste0("singleCell_LAB/HitGlio_sample1-test/plots/18-dec-24_sct_dsb-threshold_", i),
    ggheight = 20,
    ggwidth = 20
  )
}
```












```{r Plot demux result}
Idents(seu_demux) <- "MULTI_ID"

VlnPlot(seu_demux, features = "nCount_RNA", pt.size = 0.1, log = TRUE)
```






# Subset myeloid cluster to filter out cd8 cells -------------------
```{r subset myelo}
DefaultAssay(seu_singlet) <- "SCT"
Idents(seu_singlet) <- "cluster.sct"

seu_singlet_myelo <- subset(seu_singlet, idents = c(2, 6, 10))
seu_singlet_myelo

seu_singlet_myelo <- seu_normalize_var_scale_pca(
  seu_singlet_myelo,
  normalize = "sct",
  cluster.name = "cluster.pca.myelo",
  run_pca = TRUE,
  pca.reduction.name = "pca.myelo.sct",
  umap.reduction.name = "umap.myelo.pca.sct",
  dims = 1:5,
  k.param = 10,
  algorithm = 4,
  resolution = 0.4,
  group.singletons = FALSE
)
```

```{r viz isolated myelo clusters}
viz_quality(seu_singlet_myelo,
  assay = "SCT", q_features = vars,
  reduction = "umap.myelo.pca.sct", gr_by_features = gr_by_features
)

seu_dittoDotPlot(
  seu_singlet_myelo,
  vars = vars,
  group.by = "cluster.pca.myelo"
)

DefaultAssay(seu_singlet) <- "ADT"
viz_quality(seu_singlet_myelo,
  assay = "ADT", q_features = adt_features,
  reduction = "umap.myelo.pca.sct", gr_by_features = NULL
)
```

# Remove selected cells from seu_singlet (original) clusters
```{r remove cells}
Idents(seu_singlet_myelo) <- "cluster.pca.myelo"
clusters_to_remove <- subset(seu_singlet_myelo, idents = c(10, 11, 12))
clusters_to_remove

seu_singlet <- subset(
  seu_singlet,
  cells = setdiff(Cells(seu_singlet), Cells(clusters_to_remove))
)
seu_singlet
```

# Rereun normalize after removing CD8 cells from myelo clusters ------------
```{r normalize after removing cd8 from myelo clusters}
seu_singlet <- seu_normalize_var_scale_pca(
  seu_singlet,
  normalize = "sct",
  cluster.name = "cluster.sct",
  run_pca = TRUE,
  pca.reduction.name = "pca.sct",
  umap.reduction.name = "umap.sct",
  dims = 1:30,
  k.param = 20,
  algorithm = 4,
  resolution = 0.5,
  group.singletons = TRUE
)
seu_singlet <- NormalizeData(
  seu_singlet,
  assay = "ADT",
  normalization.method = "CLR"
)
```

```{r viz seu_singlet after removing cd8 from myelo clusters}
viz_quality(seu_singlet,
  assay = "SCT", q_features = vars,
  reduction = "umap.sct", gr_by_features = gr_by_features
)

seu_dittoDotPlot(
  seu_singlet,
  vars = vars,
  group.by = "cluster.sct"
)

DefaultAssay(seu_singlet) <- "ADT"
viz_quality(seu_singlet,
  assay = "ADT", q_features = adt_features,
  reduction = "umap.sct", gr_by_features = NULL
)
```

Decided to first remove cd8 from myeloid clusters based on sct/pca clustering, and than run totalVI analysis with data cleaned up as much as possible.

Now I want to merge it with dev, and try to remove background, i.e. more cleaning, before totalVI. Not sure if I should use totalVI when protein data seems to have big backgroud. Yes, if I'll menage to remove it, but for now I did not succeded.

# THIS IS NOT FINISHED FROM HERE, THIS IS OLD CODE


# Export SCT and ADT as anndata for python TotalVI
This code should be run independently of the pipeline
```{r totalVI model training of doublet-removed seu_singlet, eval = FALSE}
DefaultAssay(seu_singlet) <- "SCT"

## /for the purpose of scVI extract 3k variable genes (stored in scale.data)
## /and add it to the seurat object (TRUE/FALSE column)
## / ! this code is not working
# scaled_matrix <- GetAssayData(seu_singlet, layer = "scale.data", assay = "SCT")
# test <- rownames(seu_singlet) %in% rownames(scaled_matrix)
# seu_singlet <- AddMetaData(seu_singlet,
#   metadata = test,
#   col.name = "sct.var.genes"
# )
# seu_singlet@meta.data$sct.var.genes

# / trying to success, with chat (no success)
# sct_vst_counts <- seu_singlet[["SCT"]]@data
# variable_features <- VariableFeatures(seu_singlet)
# variable_features
# vst_counts_variable <- sct_vst_counts[variable_features, ]
# dim(vst_counts_variable)

## / v3 problem, not relevant to SCT normalization
# seu_singlet <- readRDS("RDS/seu_singlet-clustered-27nov2024.rds")
# seu_singlet[["RNA3"]] <- as(object = seu_singlet[["RNA"]], Class = "Assay")
# DefaultAssay(seu_singlet) <- "RNA3"
# seu_singlet[["RNA"]] <- NULL
# seu_singlet <- RenameAssays(object = seu_singlet, RNA3 = "RNA")

# ! / using all genes for now, as I don't know how to exctract info for sct
seu_singlet
adata <- convertFormat(
  seu_singlet,
  from = "seurat",
  to = "anndata",
  main_layer = "counts",
  # assay = "SCT",
  assay = "ADT",
  drop_single_values = FALSE,
  outFile = "RDS/seu_singlet_adt_06-dec-24.h5ad"
)
adata
```

# After totalVI run, bring back calculated latent
```{r Go to python and bring back latent.csv, eval=FALSE}
latent <- read.csv("/mnt/sda4/singleCell_LAB/HitGlio_sample1-test/latents/latent-seu_singlet_adt_06-dec-24.csv")

latent_mtx <- as.matrix(latent)
rownames(latent_mtx) <- colnames(seu_singlet)
latent_mtx <- latent_mtx[, -1] # remove 1st col
dim(latent_mtx)
colnames(latent_mtx) <- paste0("totalvi_", 1:20)
colnames(latent_mtx)

DefaultAssay(seu_singlet) <- "SCT"
seu_singlet
colnames(seu_singlet@meta.data)

seu_singlet[["totalvi.sct"]] <- CreateDimReducObject(
  embeddings = latent_mtx,
  key = "totalvi_",
  assay = DefaultAssay(seu_singlet)
)
seu_singlet
colnames(seu_singlet@meta.data)
```

/clustering po dodaniu latent
/ nie normalizuje bo nie trzeba i nie robie pca poniewaz używam totalvi
```{r clustering after adding latent, eval = FALSE}
seu_singlet <- seu_normalize_var_scale_pca(
  seu_singlet,
  normalize = FALSE,
  cluster.name = "cluster.totalvi",
  run_pca = FALSE, # not runnign pca, changing clusters in totalvi reduction
  pca.reduction.name = "totalvi.sct", # instead of pca
  umap.reduction.name = "umap.totalvi.sct",
  dims = 1:20,
  k.param = 20,
  algorithm = 4,
  resolution = 0.1,
  group.singletons = FALSE
)
```

# INSPECT features (mainly for CD8) after totalVI latent SCT clustering
```{r inspect, eval = FALSE}
DefaultAssay(seu_singlet) <- "SCT"

vars <- c("CD8A", "CD8B")
gr_by_features <- c("MULTI_ID", "cluster.totalvi")

viz_quality(seu_singlet,
  assay = "SCT", q_features = vars,
  reduction = "umap.totalvi.sct", gr_by_features = gr_by_features
)

seu_dittoDotPlot(
  seu_singlet,
  vars = vars,
  group.by = "cluster.totalvi"
)

DefaultAssay(seu_singlet) <- "ADT"
adt_features <- c("TotalSeqC-CD8", "TotalSeqC-CD4")

viz_quality(seu_singlet,
  assay = "ADT", q_features = adt_features,
  reduction = "umap.totalvi.sct", gr_by_features = NULL
)
```



# 15. Rerun totalVI after removing CD8 cells from myeloid cluster -----
## 15.1 convert to anndata
seu_singlet
adata <- convertFormat(
  seu_singlet,
  from = "seurat",
  to = "anndata",
  main_layer = "counts",
  assay = "SCT",
  # assay = "ADT",
  drop_single_values = FALSE,
  outFile = "RDS/seu_singlet_sct_wo-cd8inMyelo_06dec24_adt-quant99.h5ad"
)

## 15.2 Go to python, and bring back latent.csv --------------------
latent <- read.csv("latents/latent-seu_singlet_sct_wo-cd8inMyelo_06dec24_adt-quant99.csv")

latent_mtx <- as.matrix(latent)
rownames(latent_mtx) <- colnames(seu_singlet)
latent_mtx <- latent_mtx[, -1] # remove 1st col
dim(latent_mtx)
colnames(latent_mtx) <- paste0("totalvi_", 1:20)
colnames(latent_mtx)

DefaultAssay(seu_singlet) <- "SCT"
seu_singlet
colnames(seu_singlet@meta.data)

seu_singlet[["totalvi.sct"]] <- CreateDimReducObject(
  embeddings = latent_mtx,
  key = "totalvi_",
  assay = DefaultAssay(seu_singlet)
)
seu_singlet
colnames(seu_singlet@meta.data)

# 16. Clustering po dodaniu latent-wo-myelo -------------------------
# / nie normalizuje bo nie trzeba i nie robie pca poniewaz używam totalvi
seu_singlet <- seu_normalize_var_scale_pca(
  seu_singlet,
  normalize = FALSE,
  cluster.name = "cluster.totalvi",
  run_pca = FALSE, # not runnign pca, changing clusters in totalvi reduction
  pca.reduction.name = "totalvi.sct", # instead of pca
  umap.reduction.name = "umap.totalvi.sct",
  dims = 1:20,
  k.param = 20,
  algorithm = 4,
  resolution = 0.8,
  # resolution = 1.2,
  group.singletons = FALSE
)

# 17. INSPECT features for CD8 after removing from myelo ---------
DefaultAssay(seu_singlet) <- "SCT"
inferno <- viridis::scale_color_viridis(option = "inferno")

vars <- t_cell_markers
vars <- macrophages
vars <- ein_like
vars <- inpc_like
vars <- c("CD8A", "CD8B")
vars <- c("GFAP", "ALDH1L1", "AQP4", "SLC1A2", "CD44", "S100B", "VIM") # astrocytes
vars <- c("GFAP", "AQP4", "S100B", "VIM")

vars <- c(
  "CST3", "S100B", "SLC1A3", "HEPN1", "HOPX", "MT3", "SPARCL1", "MLC1", "GFAP",
  "FABP7", "BCAN", "PON2", "METTL7B", "SPARC", "GATM", "RAMP1", "PMP2", "AQP4",
  "DBI", "EDNRB", "PTPRZ1", "CLU", "PMP22", "ATP1A2", "S100A16", "HEY1", "PCDHGC3", "TTYH1", "NDRG2", "PRCP", "ATP1B2", "AGT", "PLTP", "GPM6B", "F3", "RAB31", "PPAP2B", "ANXA5", "TSPAN7"
) # AC-like

vars <- c(
  "CSTB", "ADM", "BNIP3", "BNIP3L", "ENO2", "FAM162A", "RALA", "CYTIP", "ADAM8",
  "C15orf48", "CD109", "ENO1", "SCD", "PLP2", "RAB42", "S100A10", "S100A6", "HK2",
  "SLC2A1", "CXCL8", "LGALS1", "TIMP1", "PLIN2", "CTSL", "LDHA", "NDRG1", "HILPDA",
  "ERO1A", "NUPR1", "MT2A"
)

vars <- c("TGFBI", "CD14", "CD163", "SELENOP", "GPNMB")

vars <- c("RPL13", "RPL19", "RPS15", "RPL28", "RPLP1", "RPL35")
vars <- c("CD83", "CD74", "HLA-DRA", "HLA-DQA1", "DDIT3", "THBS1", "CD69", "FOSB", "TNFAIP3", "ZFP36", "BAG3")

vars <- c("GFAP", "PTPRZ1", "SCG3", "C1orf61", "SOX2", "SOX6", "BCAN", "NOV/AN1", "NRCAM", "TUBB2B")

seu_FeaturePlot(
  seu_singlet,
  assay = "SCT",
  features = vars,
  reduction = "umap.totalvi.sct",
  # reduction = "umap.sct",
  color = inferno,
  max.cutoff = NA,
  show = FALSE,
  save_path = "plots/temp",
  ggheight = 30,
  ggwidth = 30
)

seu_DimPlot(
  seu_singlet,
  reduction = "umap.totalvi.sct",
  # reduction = "umap.sct",
  # group.by = c("MULTI_ID", "cluster.totalvi"),
  group.by = c("my_id"),
  # group.by = c("MULTI_ID", "cluster.sct"),
  show = FALSE,
  save_path = "plots/temp",
  ggheight = 7,
  ggwidth = 10
) & NoLegend()

seu_dittoDotPlot(
  seu_singlet,
  vars = vars,
  group.by = "seurat_clusters"
)

# 18. Vizualize ADT features after all that cleanup -------------
DefaultAssay(seu_singlet) <- "ADT"
features <- Features(seu_singlet)
length(features)
f1 <- features[1:10]
f2 <- features[11:21]
f3 <- features[22:32]
f4 <- features[33:42]
f_list <- list(f1, f2, f3, f4)

for (i in seq_along(1:4)) {
  seu_FeaturePlot(
    seu_singlet, "ADT",
    reduction = "umap.totalvi.sct",
    color = inferno,
    features = f_list[[i]],
    max.cutoff = NA,
    show = FALSE,
    save_path = paste0("plots/06-dec-24_scv-totalvi-final/adt-panel_", i),
    ggheight = 20,
    ggwidth = 20
  )
}

saveRDS(seu_singlet, "RDS/06-dec-24_seu_singlet_final_doublet-500_wo-cd8-myelo_sct_totalvi.rds")


# viz. proteins
features <- paste0("TotalSeqC-", c(
  "CD86", "CD172a", "CD184", "CD59", "CD11b", "CD11c", "CD49d", "CD192", "CD366", "CD74", "CD52", "CD14", "HLA-DR", "CD45"
))
features
seu_FeaturePlot(
  seu_singlet_myelo_dc, "ADT",
  reduction = "umap.totalvi.sct",
  color = inferno,
  features = features,
  max.cutoff = NA,
  show = FALSE,
  label = FALSE,
  save_path = "plots/temp.png",
  ggheight = 15,
  ggwidth = 15
)







# 18. AZIMUTH
library(Azimuth)

DefaultAssay(seu_singlet) <- "SCT"

seu_singlet_azi <- RunAzimuth(
  seu_singlet,
  reference = "humancortexref",
  do.adt = TRUE,
  assay = "SCT",
  query.modality = "SCT"
)

colnames(seu_singlet_azi@meta.data)

seu_singlet_azi <- NormalizeData(seu_singlet_azi)

Idents(seu_singlet_azi) <- "predicted.celltype.l2"

seu_DimPlot(
  seu_singlet_azi,
  reduction = "umap.totalvi.sct",
  group.by = c("cluster.totalvi", "predicted.subclass"),
  # group.by = c("cluster.totalvi", "predicted.celltype.l2"),
  show = TRUE,
  # save_path = "plots/scv-500-doublets/PBMCref",
  ggheight = NA,
  ggwidth = NA
) & NoLegend()




# 19. Identify myeloid clusters ------------
# / subset potentially myeloid/dc clusters
Idents(seu_singlet_azi) <- "cluster.totalvi"
seu_singlet_myelo_dc <- subset(seu_singlet_azi, idents = c(4, 7, 6, 12, 20, 5, 24, 21, 22, 10))

Idents(seu_singlet_azi) <- "my_id"
seu_DimPlot(
  seu_singlet_myelo_dc,
  reduction = "umap.totalvi.sct",
  # group.by = c("cluster.totalvi"),
  group.by = c("my_id"),
  show = TRUE,
  save_path = "plots/temp",
  ggheight = 6,
  ggwidth = 6
)
Idents(seu_singlet_myelo_dc) <- "my_id"
Idents(seu_singlet_myelo_dc)

?FindMarkers
Idents(seu_singlet) <- "cluster.totalvi"
cluster_4_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 4,
  min.pct = 0.25
)
rownames(cluster_4_markers)[1:100]

cluster_7_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 7,
  min.pct = 0.25
)
rownames(cluster_7_markers)[1:100]

cluster_6_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 6,
  min.pct = 0.25
)
rownames(cluster_6_markers)[1:100]

cluster_12_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 12,
  min.pct = 0.25
)
rownames(cluster_12_markers)[1:100]

cluster_5_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 5,
  min.pct = 0.25
)
rownames(cluster_5_markers)[1:100]

cluster_10_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 10,
  min.pct = 0.25
)
rownames(cluster_10_markers)[1:100]

cluster_20_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 20,
  min.pct = 0.25
)
rownames(cluster_20_markers)[1:100]

cluster_22_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 22,
  min.pct = 0.25
)
rownames(cluster_22_markers)[1:100]

cluster_21_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 21,
  min.pct = 0.25
)
rownames(cluster_21_markers)[1:100]

cluster_24_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 24,
  min.pct = 0.25
)
rownames(cluster_24_markers)[1:100]



# identify lymphoid clusters -------------------
Idents(seu_singlet) <- "my_id"

seu_DimPlot(
  seu_singlet_lymp,
  reduction = "umap.totalvi.sct",
  # group.by = c("cluster.totalvi"),
  group.by = c("my_id"),
  show = TRUE,
  save_path = "plots/temp",
  ggheight = 8,
  ggwidth = 8
)

seu_singlet_lymp <- subset(seu_singlet, idents = c(
  3, 16, 18, 14, 25, 1, 23, 2, 11, 17, 13, 8, 9, 19
))













cluster_16_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 16,
  min.pct = 0.25
)

rownames(cluster_16_markers)[1:100]

cluster_18_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 18,
  min.pct = 0.25
)
rownames(cluster_18_markers)[1:100]

# 15. vizualize clonality ----------------------------------------
library(scRepertoire)

tcr_sample <- read_csv("/mnt/sdb1/runs/sample1_multilane_spec_index_NNN/outs/per_sample_outs/sample1/vdj_t/filtered_contig_annotations.csv")

dim(tcr_sample)
# 16400 31

tcr_sample_list <- list(tcr_sample)

combined_tcr <- combineTCR(
  tcr_sample_list,
  removeNA = FALSE,
  removeMulti = FALSE,
  filterMulti = FALSE
)

clonalQuant(
  combined_tcr,
  cloneCall = "strict",
  chain = "both",
  scale = TRUE
)

clonalLength(
  combined_tcr,
  cloneCall = "aa",
  chain = "both"
)

seu_singlet <- combineExpression(
  combined_tcr,
  seu_singlet,
  cloneCall = "gene",
  proportion = TRUE
)

colorblind_vector <- hcl.colors(n = 7, palette = "inferno", fixup = TRUE)

seu_DimPlot(
  seu_singlet,
  reduction = "umap.totalvi.sct",
  # group.by = c("cluster.totalvi"),
  group.by = c("cloneSize"),
  show = FALSE,
  label = FALSE,
  save_path = "plots/temp",
  ggheight = 6,
  ggwidth = 8
)


DimPlot(
  seu_singlet,
  reduction = "umap.totalvi.sct",
  group.by = "cloneSize"
) + scale_color_manual(values = rev(colorblind_vector[c(1, 3, 4, 5, 7)]))


# 16. Final Naming of clusters -------------------------------------
## copy seurat_clusters column
seu_singlet[["my_id"]] <- seu_singlet[["cluster.totalvi"]]
# seu_singlet_azi[["my_id"]] <- seu_singlet[["cluster.totalvi"]]
Idents(seu_singlet) <- "my_id"
# Idents(seu_singlet_azi) <- "my_id"
seu_singlet$my_id <- plyr::mapvalues(
  # seu_singlet_azi$my_id <- plyr::mapvalues(
  Idents(seu_singlet),
  from = c(1:25),
  to = c(
    "CD8 EM",
    "CD4 Naive",
    "NK",
    "TAM_1",
    "Astrocytes",
    "Monocytes",
    "TAM_2",
    "CD8 Naive/CD8 EM (CSF)",
    "CD4 Naive (CSF)",
    "pDC",
    "CD4 CTL EOMES",
    "cDC",
    "CD4 TREG",
    "CD8 TEMRA",
    "Prolif.",
    "Neutrophils",
    "17",
    "TAN",
    "19",
    "cDC1",
    "?",
    "B-cells",
    "23",
    "OligoDC",
    "25"
  )
)

# 3. Subset csf and tumor singlets (for tcr analysis) -------
## / this goes to the script where tcr data in analyzed/
DefaultAssay(seu_singlet) <- "SCT"
seu_singlet_csf <- subset(
  seu_singlet,
  subset = MULTI_ID == "csf"
)
seu_singlet_tumor <- subset(
  seu_singlet,
  subset = MULTI_ID == "tumor"
)

seu_singlet_tumor <- combineExpression(
  combined_tcr,
  seu_singlet_tumor,
  cloneCall = "gene",
  proportion = TRUE
)
seu_singlet_csf <- combineExpression(
  combined_tcr,
  seu_singlet_csf,
  cloneCall = "gene",
  proportion = TRUE
)
seu_tumor_tcr4 <- Run.ProjecTILs(
  seu_singlet_tumor,
  ref = ref_cd4,
  ncores = 1
)
seu_csf_tcr4 <- Run.ProjecTILs(
  seu_singlet_csf,
  ref = ref_cd4,
  ncores = 1
)
seu_tumor_tcr8 <- Run.ProjecTILs(
  seu_singlet_tumor,
  ref = ref_cd8,
  ncores = 1
)
seu_csf_tcr8 <- Run.ProjecTILs(
  seu_singlet_csf,
  ref = ref_cd8,
  ncores = 1
)




p1 <- plot.projection(ref_cd4) + theme(aspect.ratio = 1)
p2 <- plot.projection(
  ref_cd4, seu_tumor_tcr4,
  linesize = 0.3, pointsize = 0.5
) + theme(aspect.ratio = 1)
p1 | p2

ggsave("plots/seu_tumor_tcr4.png", p2)




p3 <- plot.statepred.composition(ref_cd8, seu_tumor_tcr8, metric = "Percent")
ggsave("plots/seu_tumor_tcr4_percent-composition.png", p3,
  height = 3,
  width = 5
)

p4 <- plot.states.radar(
  ref = ref_cd4, seu_tumor_tcr4, min.cells = 30
)
ggsave("plots/seu_tumor_tcr4_genes-radar.png", p4)














# 17. VIZ cycling cells (this should go much earlier)
test <- RunPCA(seu_singlet, features = c(s_genes, g2m_genes))
DimPlot(test, reduction = "pca")
DimPlot(test, reduction = "pca")

DimPlot(test,
  reduction = "pca",
  group.by = "Phase",
  split.by = "Phase"
)

FeaturePlot(
  test, "percent_ribo",
  reduction = "umap.sct",
  label = TRUE
)

Idents(seu_singlet) <- "my_id"
FeaturePlot(
  seu_singlet,
  # c("CX3CR1", "P2RY12"),
  # c("IL1B", "TNF", "CCL4"),
  # c("GFAP"),
  # c("FOLR2", "LYVE1", "MRC1", "CD206"),
  # "CD8B",
  tcels,
  # "PDCD1",
  # "CADM1",
  # c("TREM1", "TREM2"),
  # c("p53", "ATM", "CHEK2"),
  # c("HIF1A"),
  # c("BAX", "CASP3", "FAS"),
  # c("MICA", "MICB"),
  # c("TREM2", "SPP1"),
  # c("CD34", "CD38", "CD123", "CD117", "CD13", "CD33", "HLA‐DR"),
  reduction = "umap",
  label = TRUE
)

DefaultAssay(seu_singlet) <- "SCT"
colnames(seu_singlet@meta.data)
Idents(seu_singlet) <- "cluster.totalvi"
Idents(seu_singlet)
?SetIdent
cluster_microglia.markers <- FindMarkers(
  seu_singlet,
  ident.1 = 7,
  min.pct = 0.25
)
rownames(cluster_microglia.markers)[1:100]

cluster_3.markers <- FindMarkers(
  seu_singlet,
  ident.1 = "3",
  min.pct = 0.25
)
rownames(cluster_3.markers)[1:30]

saveRDS(seu_singlet, "RDS/seu_singlet-clustered-totalvi-28nov2024.rds")



FeatureScatter(seu_test, feature1 = "nFeature_RNA", feature2 = "nCount_ADT")
seu_test <- subset(seu_singlet, subset = nCount_ADT < quantile(seu_singlet$nCount_ADT, 0.99))

DefaultAssay(seu_test) <- "SCT"

seu_test <- seu_normalize_var_scale_pca(
  seu_test,
  normalize = "SCT",
  cluster.name = "cluster.pca",
  run_pca = TRUE, # not runnign pca, changing clusters in totalvi reduction
  pca.reduction.name = "pca.sct", # instead of pca
  umap.reduction.name = "umap.sct",
  dims = 1:20,
  k.param = 20,
  algorithm = 4,
  resolution = 0.8,
  group.singletons = FALSE
)
seu_test <- NormalizeData(
  seu_test,
  assay = "ADT",
  normalization.method = "CLR"
)

for (i in seq_along(1:4)) {
  seu_FeaturePlot(
    seu_test, "ADT",
    reduction = "umap.sct",
    color = inferno,
    features = f_list[[i]],
    max.cutoff = NA,
    show = FALSE,
    save_path = paste0("plots/06-dec-24_scv-totalvi-final/adt-panel_", i),
    ggheight = 20,
    ggwidth = 20
  )
}
seu_test

vars <- c("GFAP", "PTPRZ1", "SCG3", "C1orf61", "SOX2", "SOX6", "BCAN", "NOV/AN1", "NRCAM", "TUBB2B") # tumor cell

vars <- c("SOX2", "SOX6", "MK167", "TOP2A", "CDC34", "PBK", "UBE2T", "CENPF")

vars <- c("MAG", "CLDN11", "APLP1", "TMEM144", "CNP", "OPALIN", "EDIL3", "LARP6", "MOG", "AMER2", "TUBB4A")

vars <- c("MK167", "STMN1", "TYMS", "TOP2A", "TUBB")

vars <- c("P2RY12", "CX3CR1", "FCGR1A", "CH25H", "CCL4L2")

vars <- c("TGFB1", "CD14", "CD163", "SELENOP", "GPNMB")

vars <- c("ENO1", "SCD", "PLP2", "RAB42", "S100A10", "S100A6", "HK2", "SLC2A1", "CXCL8", "LGALS1", "TIMP1", "PLIN2", "CTSL", "LDHA", "NDRG1", "HILPDA", "ERO1A", "NUPR1", "MT2A")

vars <- c("SPP1", "RGS16", "FSCN1", "HAMP", "BIN1", "IBSP")

vars <- c("IFIT1", "IFI44L", "IFIT3", "IFI16", "IFI44", "FGL2", "HERC5", "IFIT2", "ISG15", "CXCL10", "MX1", "MX2")

vars <- c("APOC1", "APOE", "CD63", "LGALS3", "ACP5", "GCHFR", "OTOA", "PLAG27", "SDS")

vars <- c("CD83", "CD74", "HLA-DRA", "HLA-DQA1", "DDIT3", "THBS1", "CD69", "FOSB", "TNFAIP3", "ZFP36", "BAG3")

vars <- c("RPL13", "RPL19", "RPS15", "RPL28", "RPLP1", "RPL35")


DefaultAssay(seu_singlet)
seu_singlet

seu_singlet_myelo_dc <- Seurat::AddModuleScore(
  seu_singlet_myelo_dc,
  features = list(vars),
  name = "Mg-TAM",
  layer = scale.data
)

p4 <- Seurat::VlnPlot(
  seu_singlet_myelo_dc,
  "Mg-TAM1",
) + ggsave("plots/temp.png", p4)


colnames(seu_singlet_myelo_dc@meta.data)

# check backgroud signals
seu_singlet <- readRDS("RDS/06-dec-24_seu_singlet_final_doublet-500_wo-cd8-myelo_sct_totalvi.rds")
isotype_controls <- c(
  "TotalSeqC-IgG2a-Ctrl",
  "TotalSeqC-IgG2b-Ctrl",
  "TotalSeqC-IgG1-Ctrl",
  "TotalSeqC-IgG-Ctrl"
)
DefaultAssay(seu_singlet) <- "ADT"
Features(seu_singlet)
VlnPlot(seu_singlet, features = isotype_controls)

VlnPlot(
  seu_singlet,
  features = c("TotalSeqC-CD137", "TotalSeqC-IgG1-Ctrl", "TotalSeqC-IgG-Ctrl")
)
