---
title: "HIT-GLIO sample1 analysis #1"
author: "MF"
date: "`r Sys.Date()`"
---

according to: https://satijalab.org/seurat/articles/hashing_vignette

```{r}
library(httpgd)
httpgd::hgd_url()
# hgd()
```

rmarkdown::render("/mnt/sda4/singleCell_LAB/HitGlio_sample1-test/scripts/01_demux_filter.Rmd")

# Load Libraries
```{r Libraries, warning = FALSE, message = FALSE}
suppressPackageStartupMessages({
  library(Seurat)
  library(DropletUtils)
  library(ggplot2)
  library(Matrix)
  library(tidyverse)
  library(clustree)
  library(reticulate) # load packages required by leiden algo
  reticulate::use_condaenv("/mnt/sda4/conda/envs/R")
  library(viridis) # colors
  library(DoubletFinder)
  options(future.globals.maxSize = 3.0 * 1e9) # 3GB
  library(sceasy) # convert to anndata, scvi integration
  library(dittoSeq)
  library(dsb)
})
```


# DSB



# 18. Vizualize ADT features after all that cleanup -------------
```{r}
DefaultAssay(seu_singlet) <- "ADT.dsb"
features <- Features(seu_singlet)
length(features)
f1 <- features[1:10]
f2 <- features[11:21]
f3 <- features[22:32]
f4 <- features[33:42]
f_list <- list(f1, f2, f3, f4)

for (i in seq_along(1:4)) {
  seu_FeaturePlot(
    seu_singlet, "ADT.dsb",
    reduction = "umap.sct",
    color = inferno,
    features = f_list[[i]],
    max.cutoff = NA,
    show = FALSE,
    save_path = paste0("plots/18-dec-24_sct_dsb_2/adt-panel_", i),
    ggheight = 20,
    ggwidth = 20
  )
}
```

```{r}

```












```{r Plot demux result}
Idents(seu_demux) <- "MULTI_ID"

VlnPlot(seu_demux, features = "nCount_RNA", pt.size = 0.1, log = TRUE)
```








Decided to first remove cd8 from myeloid clusters based on sct/pca clustering, and than run totalVI analysis with data cleaned up as much as possible.

Now I want to merge it with dev, and try to remove background, i.e. more cleaning, before totalVI. Not sure if I should use totalVI when protein data seems to have big backgroud. Yes, if I'll menage to remove it, but for now I did not succeded.

# THIS IS NOT FINISHED FROM HERE, THIS IS OLD CODE


# Export SCT and ADT as anndata for python TotalVI
This code should be run independently of the pipeline
```{r totalVI model training of doublet-removed seu_singlet, eval = FALSE}
DefaultAssay(seu_singlet) <- "SCT"

## /for the purpose of scVI extract 3k variable genes (stored in scale.data)
## /and add it to the seurat object (TRUE/FALSE column)
## / ! this code is not working
# scaled_matrix <- GetAssayData(seu_singlet, layer = "scale.data", assay = "SCT")
# test <- rownames(seu_singlet) %in% rownames(scaled_matrix)
# seu_singlet <- AddMetaData(seu_singlet,
#   metadata = test,
#   col.name = "sct.var.genes"
# )
# seu_singlet@meta.data$sct.var.genes

# / trying to success, with chat (no success)
# sct_vst_counts <- seu_singlet[["SCT"]]@data
# variable_features <- VariableFeatures(seu_singlet)
# variable_features
# vst_counts_variable <- sct_vst_counts[variable_features, ]
# dim(vst_counts_variable)

## / v3 problem, not relevant to SCT normalization
# seu_singlet <- readRDS("RDS/seu_singlet-clustered-27nov2024.rds")
# seu_singlet[["RNA3"]] <- as(object = seu_singlet[["RNA"]], Class = "Assay")
# DefaultAssay(seu_singlet) <- "RNA3"
# seu_singlet[["RNA"]] <- NULL
# seu_singlet <- RenameAssays(object = seu_singlet, RNA3 = "RNA")

# ! / using all genes for now, as I don't know how to exctract info for sct
seu_singlet
adata <- convertFormat(
  seu_singlet,
  from = "seurat",
  to = "anndata",
  main_layer = "counts",
  # assay = "SCT",
  assay = "ADT",
  drop_single_values = FALSE,
  outFile = "RDS/seu_singlet_adt_06-dec-24.h5ad"
)
adata
```

# After totalVI run, bring back calculated latent
```{r Go to python and bring back latent.csv, eval=FALSE}
latent <- read.csv("/mnt/sda4/singleCell_LAB/HitGlio_sample1-test/latents/latent-seu_singlet_adt_06-dec-24.csv")

latent_mtx <- as.matrix(latent)
rownames(latent_mtx) <- colnames(seu_singlet)
latent_mtx <- latent_mtx[, -1] # remove 1st col
dim(latent_mtx)
colnames(latent_mtx) <- paste0("totalvi_", 1:20)
colnames(latent_mtx)

DefaultAssay(seu_singlet) <- "SCT"
seu_singlet
colnames(seu_singlet@meta.data)

seu_singlet[["totalvi.sct"]] <- CreateDimReducObject(
  embeddings = latent_mtx,
  key = "totalvi_",
  assay = DefaultAssay(seu_singlet)
)
seu_singlet
colnames(seu_singlet@meta.data)
```

/clustering po dodaniu latent
/ nie normalizuje bo nie trzeba i nie robie pca poniewaz używam totalvi
```{r clustering after adding latent, eval = FALSE}
seu_singlet <- seu_normalize_var_scale_pca(
  seu_singlet,
  normalize = FALSE,
  cluster.name = "cluster.totalvi",
  run_pca = FALSE, # not runnign pca, changing clusters in totalvi reduction
  pca.reduction.name = "totalvi.sct", # instead of pca
  umap.reduction.name = "umap.totalvi.sct",
  dims = 1:20,
  k.param = 20,
  algorithm = 4,
  resolution = 0.1,
  group.singletons = FALSE
)
```

# INSPECT features (mainly for CD8) after totalVI latent SCT clustering
```{r inspect, eval = FALSE}
DefaultAssay(seu_singlet) <- "SCT"

vars <- c("CD8A", "CD8B")
gr_by_features <- c("MULTI_ID", "cluster.totalvi")

viz_quality(seu_singlet,
  assay = "SCT", q_features = vars,
  reduction = "umap.totalvi.sct", gr_by_features = gr_by_features
)

seu_dittoDotPlot(
  seu_singlet,
  vars = vars,
  group.by = "cluster.totalvi"
)

DefaultAssay(seu_singlet) <- "ADT"
adt_features <- c("TotalSeqC-CD8", "TotalSeqC-CD4")

viz_quality(seu_singlet,
  assay = "ADT", q_features = adt_features,
  reduction = "umap.totalvi.sct", gr_by_features = NULL
)
```



# 15. Rerun totalVI after removing CD8 cells from myeloid cluster -----
## 15.1 convert to anndata
seu_singlet
adata <- convertFormat(
  seu_singlet,
  from = "seurat",
  to = "anndata",
  main_layer = "counts",
  assay = "SCT",
  # assay = "ADT",
  drop_single_values = FALSE,
  outFile = "RDS/seu_singlet_sct_wo-cd8inMyelo_06dec24_adt-quant99.h5ad"
)

## 15.2 Go to python, and bring back latent.csv --------------------
latent <- read.csv("latents/latent-seu_singlet_sct_wo-cd8inMyelo_06dec24_adt-quant99.csv")

latent_mtx <- as.matrix(latent)
rownames(latent_mtx) <- colnames(seu_singlet)
latent_mtx <- latent_mtx[, -1] # remove 1st col
dim(latent_mtx)
colnames(latent_mtx) <- paste0("totalvi_", 1:20)
colnames(latent_mtx)

DefaultAssay(seu_singlet) <- "SCT"
seu_singlet
colnames(seu_singlet@meta.data)

seu_singlet[["totalvi.sct"]] <- CreateDimReducObject(
  embeddings = latent_mtx,
  key = "totalvi_",
  assay = DefaultAssay(seu_singlet)
)
seu_singlet
colnames(seu_singlet@meta.data)

# 16. Clustering po dodaniu latent-wo-myelo -------------------------
# / nie normalizuje bo nie trzeba i nie robie pca poniewaz używam totalvi
seu_singlet <- seu_normalize_var_scale_pca(
  seu_singlet,
  normalize = FALSE,
  cluster.name = "cluster.totalvi",
  run_pca = FALSE, # not runnign pca, changing clusters in totalvi reduction
  pca.reduction.name = "totalvi.sct", # instead of pca
  umap.reduction.name = "umap.totalvi.sct",
  dims = 1:20,
  k.param = 20,
  algorithm = 4,
  resolution = 0.8,
  # resolution = 1.2,
  group.singletons = FALSE
)

# 17. INSPECT features for CD8 after removing from myelo ---------
DefaultAssay(seu_singlet) <- "SCT"
inferno <- viridis::scale_color_viridis(option = "inferno")

vars <- t_cell_markers
vars <- macrophages
vars <- ein_like
vars <- inpc_like
vars <- c("CD8A", "CD8B")
vars <- c("GFAP", "ALDH1L1", "AQP4", "SLC1A2", "CD44", "S100B", "VIM") # astrocytes
vars <- c("GFAP", "AQP4", "S100B", "VIM")

vars <- c(
  "CST3", "S100B", "SLC1A3", "HEPN1", "HOPX", "MT3", "SPARCL1", "MLC1", "GFAP",
  "FABP7", "BCAN", "PON2", "METTL7B", "SPARC", "GATM", "RAMP1", "PMP2", "AQP4",
  "DBI", "EDNRB", "PTPRZ1", "CLU", "PMP22", "ATP1A2", "S100A16", "HEY1", "PCDHGC3", "TTYH1", "NDRG2", "PRCP", "ATP1B2", "AGT", "PLTP", "GPM6B", "F3", "RAB31", "PPAP2B", "ANXA5", "TSPAN7"
) # AC-like

vars <- c(
  "CSTB", "ADM", "BNIP3", "BNIP3L", "ENO2", "FAM162A", "RALA", "CYTIP", "ADAM8",
  "C15orf48", "CD109", "ENO1", "SCD", "PLP2", "RAB42", "S100A10", "S100A6", "HK2",
  "SLC2A1", "CXCL8", "LGALS1", "TIMP1", "PLIN2", "CTSL", "LDHA", "NDRG1", "HILPDA",
  "ERO1A", "NUPR1", "MT2A"
)

vars <- c("TGFBI", "CD14", "CD163", "SELENOP", "GPNMB")

vars <- c("RPL13", "RPL19", "RPS15", "RPL28", "RPLP1", "RPL35")
vars <- c("CD83", "CD74", "HLA-DRA", "HLA-DQA1", "DDIT3", "THBS1", "CD69", "FOSB", "TNFAIP3", "ZFP36", "BAG3")

vars <- c("GFAP", "PTPRZ1", "SCG3", "C1orf61", "SOX2", "SOX6", "BCAN", "NOV/AN1", "NRCAM", "TUBB2B")

seu_FeaturePlot(
  seu_singlet,
  assay = "SCT",
  features = vars,
  reduction = "umap.totalvi.sct",
  # reduction = "umap.sct",
  color = inferno,
  max.cutoff = NA,
  show = FALSE,
  save_path = "plots/temp",
  ggheight = 30,
  ggwidth = 30
)

seu_DimPlot(
  seu_singlet,
  reduction = "umap.totalvi.sct",
  # reduction = "umap.sct",
  # group.by = c("MULTI_ID", "cluster.totalvi"),
  group.by = c("my_id"),
  # group.by = c("MULTI_ID", "cluster.sct"),
  show = FALSE,
  save_path = "plots/temp",
  ggheight = 7,
  ggwidth = 10
) & NoLegend()

seu_dittoDotPlot(
  seu_singlet,
  vars = vars,
  group.by = "seurat_clusters"
)

# 18. Vizualize ADT features after all that cleanup -------------
DefaultAssay(seu_singlet) <- "ADT"
features <- Features(seu_singlet)
length(features)
f1 <- features[1:10]
f2 <- features[11:21]
f3 <- features[22:32]
f4 <- features[33:42]
f_list <- list(f1, f2, f3, f4)

for (i in seq_along(1:4)) {
  seu_FeaturePlot(
    seu_singlet, "ADT",
    reduction = "umap.totalvi.sct",
    color = inferno,
    features = f_list[[i]],
    max.cutoff = NA,
    show = FALSE,
    save_path = paste0("plots/06-dec-24_scv-totalvi-final/adt-panel_", i),
    ggheight = 20,
    ggwidth = 20
  )
}

saveRDS(seu_singlet, "RDS/06-dec-24_seu_singlet_final_doublet-500_wo-cd8-myelo_sct_totalvi.rds")


# viz. proteins
features <- paste0("TotalSeqC-", c(
  "CD86", "CD172a", "CD184", "CD59", "CD11b", "CD11c", "CD49d", "CD192", "CD366", "CD74", "CD52", "CD14", "HLA-DR", "CD45"
))
features
seu_FeaturePlot(
  seu_singlet_myelo_dc, "ADT",
  reduction = "umap.totalvi.sct",
  color = inferno,
  features = features,
  max.cutoff = NA,
  show = FALSE,
  label = FALSE,
  save_path = "plots/temp.png",
  ggheight = 15,
  ggwidth = 15
)







# 18. AZIMUTH
library(Azimuth)

DefaultAssay(seu_singlet) <- "SCT"

seu_singlet_azi <- RunAzimuth(
  seu_singlet,
  reference = "humancortexref",
  do.adt = TRUE,
  assay = "SCT",
  query.modality = "SCT"
)

colnames(seu_singlet_azi@meta.data)

seu_singlet_azi <- NormalizeData(seu_singlet_azi)

Idents(seu_singlet_azi) <- "predicted.celltype.l2"

seu_DimPlot(
  seu_singlet_azi,
  reduction = "umap.totalvi.sct",
  group.by = c("cluster.totalvi", "predicted.subclass"),
  # group.by = c("cluster.totalvi", "predicted.celltype.l2"),
  show = TRUE,
  # save_path = "plots/scv-500-doublets/PBMCref",
  ggheight = NA,
  ggwidth = NA
) & NoLegend()




# 19. Identify myeloid clusters ------------
# / subset potentially myeloid/dc clusters
Idents(seu_singlet_azi) <- "cluster.totalvi"
seu_singlet_myelo_dc <- subset(seu_singlet_azi, idents = c(4, 7, 6, 12, 20, 5, 24, 21, 22, 10))

Idents(seu_singlet_azi) <- "my_id"
seu_DimPlot(
  seu_singlet_myelo_dc,
  reduction = "umap.totalvi.sct",
  # group.by = c("cluster.totalvi"),
  group.by = c("my_id"),
  show = TRUE,
  save_path = "plots/temp",
  ggheight = 6,
  ggwidth = 6
)
Idents(seu_singlet_myelo_dc) <- "my_id"
Idents(seu_singlet_myelo_dc)

?FindMarkers
Idents(seu_singlet) <- "cluster.totalvi"
cluster_4_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 4,
  min.pct = 0.25
)
rownames(cluster_4_markers)[1:100]

cluster_7_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 7,
  min.pct = 0.25
)
rownames(cluster_7_markers)[1:100]

cluster_6_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 6,
  min.pct = 0.25
)
rownames(cluster_6_markers)[1:100]

cluster_12_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 12,
  min.pct = 0.25
)
rownames(cluster_12_markers)[1:100]

cluster_5_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 5,
  min.pct = 0.25
)
rownames(cluster_5_markers)[1:100]

cluster_10_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 10,
  min.pct = 0.25
)
rownames(cluster_10_markers)[1:100]

cluster_20_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 20,
  min.pct = 0.25
)
rownames(cluster_20_markers)[1:100]

cluster_22_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 22,
  min.pct = 0.25
)
rownames(cluster_22_markers)[1:100]

cluster_21_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 21,
  min.pct = 0.25
)
rownames(cluster_21_markers)[1:100]

cluster_24_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 24,
  min.pct = 0.25
)
rownames(cluster_24_markers)[1:100]



# identify lymphoid clusters -------------------
Idents(seu_singlet) <- "my_id"

seu_DimPlot(
  seu_singlet_lymp,
  reduction = "umap.totalvi.sct",
  # group.by = c("cluster.totalvi"),
  group.by = c("my_id"),
  show = TRUE,
  save_path = "plots/temp",
  ggheight = 8,
  ggwidth = 8
)

seu_singlet_lymp <- subset(seu_singlet, idents = c(
  3, 16, 18, 14, 25, 1, 23, 2, 11, 17, 13, 8, 9, 19
))













cluster_16_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 16,
  min.pct = 0.25
)

rownames(cluster_16_markers)[1:100]

cluster_18_markers <- FindMarkers(
  seu_singlet,
  test_use = "MAST",
  ident.1 = 18,
  min.pct = 0.25
)
rownames(cluster_18_markers)[1:100]

# 15. vizualize clonality ----------------------------------------
library(scRepertoire)

tcr_sample <- read_csv("/mnt/sdb1/runs/sample1_multilane_spec_index_NNN/outs/per_sample_outs/sample1/vdj_t/filtered_contig_annotations.csv")

dim(tcr_sample)
# 16400 31

tcr_sample_list <- list(tcr_sample)

combined_tcr <- combineTCR(
  tcr_sample_list,
  removeNA = FALSE,
  removeMulti = FALSE,
  filterMulti = FALSE
)

clonalQuant(
  combined_tcr,
  cloneCall = "strict",
  chain = "both",
  scale = TRUE
)

clonalLength(
  combined_tcr,
  cloneCall = "aa",
  chain = "both"
)

seu_singlet <- combineExpression(
  combined_tcr,
  seu_singlet,
  cloneCall = "gene",
  proportion = TRUE
)

colorblind_vector <- hcl.colors(n = 7, palette = "inferno", fixup = TRUE)

seu_DimPlot(
  seu_singlet,
  reduction = "umap.totalvi.sct",
  # group.by = c("cluster.totalvi"),
  group.by = c("cloneSize"),
  show = FALSE,
  label = FALSE,
  save_path = "plots/temp",
  ggheight = 6,
  ggwidth = 8
)


DimPlot(
  seu_singlet,
  reduction = "umap.totalvi.sct",
  group.by = "cloneSize"
) + scale_color_manual(values = rev(colorblind_vector[c(1, 3, 4, 5, 7)]))


# 16. Final Naming of clusters -------------------------------------
## copy seurat_clusters column
seu_singlet[["my_id"]] <- seu_singlet[["cluster.totalvi"]]
# seu_singlet_azi[["my_id"]] <- seu_singlet[["cluster.totalvi"]]
Idents(seu_singlet) <- "my_id"
# Idents(seu_singlet_azi) <- "my_id"
seu_singlet$my_id <- plyr::mapvalues(
  # seu_singlet_azi$my_id <- plyr::mapvalues(
  Idents(seu_singlet),
  from = c(1:25),
  to = c(
    "CD8 EM",
    "CD4 Naive",
    "NK",
    "TAM_1",
    "Astrocytes",
    "Monocytes",
    "TAM_2",
    "CD8 Naive/CD8 EM (CSF)",
    "CD4 Naive (CSF)",
    "pDC",
    "CD4 CTL EOMES",
    "cDC",
    "CD4 TREG",
    "CD8 TEMRA",
    "Prolif.",
    "Neutrophils",
    "17",
    "TAN",
    "19",
    "cDC1",
    "?",
    "B-cells",
    "23",
    "OligoDC",
    "25"
  )
)

# 3. Subset csf and tumor singlets (for tcr analysis) -------
## / this goes to the script where tcr data in analyzed/
DefaultAssay(seu_singlet) <- "SCT"
seu_singlet_csf <- subset(
  seu_singlet_lymp,
  subset = MULTI_ID == "csf"
)
seu_singlet_tumor <- subset(
  seu_singlet_lymp,
  subset = MULTI_ID == "tumor"
)

seu_singlet_tumor <- combineExpression(
  combined_tcr,
  seu_singlet_tumor,
  cloneCall = "gene",
  proportion = TRUE
)
seu_singlet_csf <- combineExpression(
  combined_tcr,
  seu_singlet_csf,
  cloneCall = "gene",
  proportion = TRUE
)
seu_tumor_tcr4 <- Run.ProjecTILs(
  seu_singlet_tumor,
  ref = ref_cd4,
  ncores = 1
)
seu_csf_tcr4 <- Run.ProjecTILs(
  seu_singlet_csf,
  ref = ref_cd4,
  ncores = 1
)
seu_tumor_tcr8 <- Run.ProjecTILs(
  seu_singlet_tumor,
  ref = ref_cd8,
  ncores = 1
)
seu_csf_tcr8 <- Run.ProjecTILs(
  seu_singlet_csf,
  ref = ref_cd8,
  ncores = 1
)




p1 <- plot.projection(ref_cd4) + theme(aspect.ratio = 1)
p2 <- plot.projection(
  ref_cd4, seu_tumor_tcr4,
  linesize = 0.3, pointsize = 0.5
) + theme(aspect.ratio = 1)
p1 | p2

ggsave("plots/seu_tumor_tcr4.png", p2)




p3 <- plot.statepred.composition(ref_cd8, seu_tumor_tcr8, metric = "Percent")
ggsave("plots/seu_tumor_tcr4_percent-composition.png", p3,
  height = 3,
  width = 5
)

p4 <- plot.states.radar(
  ref = ref_cd4, seu_tumor_tcr4, min.cells = 30
)
ggsave("plots/seu_tumor_tcr4_genes-radar.png", p4)














# 17. VIZ cycling cells (this should go much earlier)
test <- RunPCA(seu_singlet, features = c(s_genes, g2m_genes))
DimPlot(test, reduction = "pca")
DimPlot(test, reduction = "pca")

DimPlot(test,
  reduction = "pca",
  group.by = "Phase",
  split.by = "Phase"
)

FeaturePlot(
  test, "percent_ribo",
  reduction = "umap.sct",
  label = TRUE
)

Idents(seu_singlet) <- "my_id"
FeaturePlot(
  seu_singlet,
  # c("CX3CR1", "P2RY12"),
  # c("IL1B", "TNF", "CCL4"),
  # c("GFAP"),
  # c("FOLR2", "LYVE1", "MRC1", "CD206"),
  # "CD8B",
  tcels,
  # "PDCD1",
  # "CADM1",
  # c("TREM1", "TREM2"),
  # c("p53", "ATM", "CHEK2"),
  # c("HIF1A"),
  # c("BAX", "CASP3", "FAS"),
  # c("MICA", "MICB"),
  # c("TREM2", "SPP1"),
  # c("CD34", "CD38", "CD123", "CD117", "CD13", "CD33", "HLA‐DR"),
  reduction = "umap",
  label = TRUE
)

DefaultAssay(seu_singlet) <- "SCT"
colnames(seu_singlet@meta.data)
Idents(seu_singlet) <- "cluster.totalvi"
Idents(seu_singlet)
?SetIdent
cluster_microglia.markers <- FindMarkers(
  seu_singlet,
  ident.1 = 7,
  min.pct = 0.25
)
rownames(cluster_microglia.markers)[1:100]

cluster_3.markers <- FindMarkers(
  seu_singlet,
  ident.1 = "3",
  min.pct = 0.25
)
rownames(cluster_3.markers)[1:30]

saveRDS(seu_singlet, "RDS/seu_singlet-clustered-totalvi-28nov2024.rds")



FeatureScatter(seu_test, feature1 = "nFeature_RNA", feature2 = "nCount_ADT")
seu_test <- subset(seu_singlet, subset = nCount_ADT < quantile(seu_singlet$nCount_ADT, 0.99))

DefaultAssay(seu_test) <- "SCT"

seu_test <- seu_normalize_var_scale_pca(
  seu_test,
  normalize = "SCT",
  cluster.name = "cluster.pca",
  run_pca = TRUE, # not runnign pca, changing clusters in totalvi reduction
  pca.reduction.name = "pca.sct", # instead of pca
  umap.reduction.name = "umap.sct",
  dims = 1:20,
  k.param = 20,
  algorithm = 4,
  resolution = 0.8,
  group.singletons = FALSE
)
seu_test <- NormalizeData(
  seu_test,
  assay = "ADT",
  normalization.method = "CLR"
)

for (i in seq_along(1:4)) {
  seu_FeaturePlot(
    seu_test, "ADT",
    reduction = "umap.sct",
    color = inferno,
    features = f_list[[i]],
    max.cutoff = NA,
    show = FALSE,
    save_path = paste0("plots/06-dec-24_scv-totalvi-final/adt-panel_", i),
    ggheight = 20,
    ggwidth = 20
  )
}
seu_test

vars <- c("GFAP", "PTPRZ1", "SCG3", "C1orf61", "SOX2", "SOX6", "BCAN", "NOV/AN1", "NRCAM", "TUBB2B") # tumor cell

vars <- c("SOX2", "SOX6", "MK167", "TOP2A", "CDC34", "PBK", "UBE2T", "CENPF")

vars <- c("MAG", "CLDN11", "APLP1", "TMEM144", "CNP", "OPALIN", "EDIL3", "LARP6", "MOG", "AMER2", "TUBB4A")

vars <- c("MK167", "STMN1", "TYMS", "TOP2A", "TUBB")

vars <- c("P2RY12", "CX3CR1", "FCGR1A", "CH25H", "CCL4L2")

vars <- c("TGFB1", "CD14", "CD163", "SELENOP", "GPNMB")

vars <- c("ENO1", "SCD", "PLP2", "RAB42", "S100A10", "S100A6", "HK2", "SLC2A1", "CXCL8", "LGALS1", "TIMP1", "PLIN2", "CTSL", "LDHA", "NDRG1", "HILPDA", "ERO1A", "NUPR1", "MT2A")

vars <- c("SPP1", "RGS16", "FSCN1", "HAMP", "BIN1", "IBSP")

vars <- c("IFIT1", "IFI44L", "IFIT3", "IFI16", "IFI44", "FGL2", "HERC5", "IFIT2", "ISG15", "CXCL10", "MX1", "MX2")

vars <- c("APOC1", "APOE", "CD63", "LGALS3", "ACP5", "GCHFR", "OTOA", "PLAG27", "SDS")

vars <- c("CD83", "CD74", "HLA-DRA", "HLA-DQA1", "DDIT3", "THBS1", "CD69", "FOSB", "TNFAIP3", "ZFP36", "BAG3")

vars <- c("RPL13", "RPL19", "RPS15", "RPL28", "RPLP1", "RPL35")


DefaultAssay(seu_singlet)
seu_singlet

seu_singlet_myelo_dc <- Seurat::AddModuleScore(
  seu_singlet_myelo_dc,
  features = list(vars),
  name = "Mg-TAM",
  layer = scale.data
)

p4 <- Seurat::VlnPlot(
  seu_singlet_myelo_dc,
  "Mg-TAM1",
) + ggsave("plots/temp.png", p4)


colnames(seu_singlet_myelo_dc@meta.data)

# check backgroud signals
seu_singlet <- readRDS("RDS/06-dec-24_seu_singlet_final_doublet-500_wo-cd8-myelo_sct_totalvi.rds")
isotype_controls <- c(
  "TotalSeqC-IgG2a-Ctrl",
  "TotalSeqC-IgG2b-Ctrl",
  "TotalSeqC-IgG1-Ctrl",
  "TotalSeqC-IgG-Ctrl"
)
DefaultAssay(seu_singlet) <- "ADT"
Features(seu_singlet)
VlnPlot(seu_singlet, features = isotype_controls)

VlnPlot(
  seu_singlet,
  features = c("TotalSeqC-CD137", "TotalSeqC-IgG1-Ctrl", "TotalSeqC-IgG-Ctrl")
)
